```{r}
library(readr)
library(dplyr)
```

#Intro
This project's goal is to find stocks that are positively correlated and then package the pairs together and make trades based on the movements of their pairs.

#Setup Code
This chunk gets a list of all the stock files in a folder that are csvs and finds the correlation between the x stocks selected and outputs that data in a data frame along with the two stocks selected for each.
```{r}
stock_correlation = data.frame(Correlation = numeric(), Stock1 = character(), Stock2 = character(), index1 = numeric(), index2 = numeric(), stringsAsFactors = FALSE) #data frame setup

files = list.files(path = "./data/", pattern = '\\.csv$', full.names = TRUE) #gets a list of all files in the data path
stock1 = read.csv(files[1])
print(stock1[1,]) #Shows row 1 of the data we use

num_stocks = 100 #Selects how many stocks used, goes by alphabetical order
iterator = 1

combine_two_stocks = function(a, b, stockNames = c(deparse(substitute(a)), deparse(substitute(b)))) { #Finds common dates where both stacks were traded 
    rr = range(intersect(a$Date, b$Date))
    a.sub = a[a$Date >= rr[1] & a$Date <= rr[2],]
    b.sub = b[b$Date >= rr[1] & b$Date <= rr[2],]
    structure(data.frame(a.sub$Date, a.sub$Adj.Close, b.sub$Adj.Close), names = c("Date", stockNames))
}
for (i in 1: num_stocks) {
    for (j in 1: num_stocks) {
        files = list.files(path = "./data/", pattern = '\\.csv$', full.names = TRUE)
        stock1 = read.csv(files[i])
        stock2 = read.csv(files[j])

        overlap = combine_two_stocks(stock1, stock2)
        
        if (files[i] != files[j]) { #Prevents the code from having it compare stocks to itself 
            name1 = substring(files[i], 8)
            name1 = substr(name1, 1, nchar(name1) - 4) #Stock name without the folder name and file extension
            name2 = substring(files[j], 8)
            name2 = substr(name2, 1, nchar(name2) - 4) #Stock name without the folder name and file extension
            
            stock_correlation[iterator, ] = list(cor(overlap$stock1, overlap$stock2), name1, name2, i, j)
            
            iterator = iterator + 1
        }
    }
}

ordered_correlation = stock_correlation[order(stock_correlation$Correlation, decreasing = TRUE), ]
reverse_ordered_correlation = stock_correlation[order(stock_correlation$Correlation, decreasing = FALSE), ]

unique_ordered_correlation <-ordered_correlation[!duplicated(ordered_correlation[, c("Correlation")]), ]
reverse_unique_ordered_correlation <-reverse_ordered_correlation[!duplicated(reverse_ordered_correlation[, c("Correlation")]), ]

print(head(unique_ordered_correlation)) #Prints the 6 highest correlated stocks
print(head(reverse_unique_ordered_correlation)) #Prints the 6 lowest correlated stocks
```



#Graph Chunk
This chunk shows three things the graph of the ratio of the differences between two stocks, the graph of the return vs the standard deviation with indicators of the current and optimal standard deviation, and the return. This figures out when to buy and sell based off a regression model it builds off of the inputted stock data using the caret package. This is more of a visual demonstration of what is happening, more will be explained in the next chunk.
```{r}
files = list.files(path = "./data/", pattern = '\\.csv$', full.names = TRUE)
stock1num = 1
stock2num = 4

stock1 = read.csv(files[stock1num]) #A
stock2 = read.csv(files[stock2num]) #AAPL

name1 = substring(files[stock1num], 8)
name1 = substr(name1, 1, nchar(name1) - 4) #Stock name without the folder name and file extension
name2 = substring(files[stock2num], 8)
name2 = substr(name2, 1, nchar(name2) - 4) #Stock name without the folder name and file extension

combine_two_stocks = function(a, b, stockNames = c(deparse(substitute(a)), deparse(substitute(b)))) {#Finds common dates where both stacks were traded 
    rr = range(intersect(a$Date, b$Date))
    a.sub = a[a$Date >= rr[1] & a$Date <= rr[2], ]
    b.sub = b[b$Date >= rr[1] & b$Date <= rr[2], ]
    structure(data.frame(a.sub$Date, a.sub$Adj.Close, b.sub$Adj.Close), names = c("Date", stockNames))
}

overlap = combine_two_stocks(stock1, stock2)
r = overlap$stock1 / overlap$stock2
k = .85 #Standard deviation

plot_ratio = function(r, k = 1, date = seq(along = r), ...) { #function that plots the ratio of correlation between stock1 and stock2
    plot(date, r, type = "l", main = paste("Ratio of pairs", name1, "and", name2, sep = " "),...)
    abline(h = c(mean(r), mean(r) + k * sd(r), mean(r) - k * sd(r)), col = c("darkgreen", rep("red", 2 * length(k))), lty = "dashed")
}
overlap$Date = as.Date(overlap$Date, format = c("%Y-%m-%d")) #Converts overlap$Date into a R readable format

find_next_position = function(ratio, startDay = 1, k = 1, m = mean(ratio), s = sd(ratio)) { #function that finds which stock to find which stock to buy or sell
    up = m + k * s
    down = m - k * s
    if (startDay > 1)
        ratio = ratio[-(1: (startDay - 1))]
    is_extreme = ratio >= up | ratio <= down
    if (!any(is_extreme))
        return (integer())
    start = which(is_extreme)[1]
    back_to_normal = if (ratio[start] > up)
            ratio[-(1: start)] <= m
    else
        ratio[-(1: start)] >= m
    end = if (any(back_to_normal))
            which(back_to_normal)[1] + start
    else
        length(ratio)
    c(start, end) + startDay - 1
}

show_position = function(days, ratios, radius = 50) { #Draws circles on graph where the ratio is optimal to buy or sell
    symbols(days, ratios, circles = rep(radius, 2), fg = c("darkgreen", "red"), add = TRUE, inches = FALSE)
}

get_positions = function(ratio, k = 1, m = mean(ratio), s = sd(ratio)) { #Finds Where the ratio is optimal to buy or sell
    when = list()
    cur = 1
    while (cur < length(ratio)) {
        temp = find_next_position(ratio, cur, k, m, s)
        if (length(temp) == 0) #quit
        break
        when[[length(when) + 1]] = temp
        if (is.na(temp[2]) || temp[2] == length(ratio))
            break
        cur = temp[2]
    }
    when
}
pos = get_positions(r, k) #Finds Where the ratio is optimal to buy or sell
plot_ratio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio") #function that plots the ratio of correlation between stock1 and stock2
invisible(lapply(pos, function(p) show_position(overlap$Date[p], r[p])))

position_profit = function(pos, stock_price_1, stock_price_2, ratio_mean = mean(stock_price_1 / stock_price_2), p = .001, by_stock = FALSE) { #Finds the profit from the purchases
    if (is.list(pos)) {
        ans = sapply(pos, position_profit,
            stock_price_1, stock_price_2, ratio_mean, p, by_stock)
        if (by_stock)
            rownames(ans) = c("A", "B", "commission")
        return (ans)
    }

    price_stock1 = stock_price_1[pos] #Gets prices
    price_stock2 = stock_price_2[pos] #Gets prices
    
    units_of_stock1 = 1 / price_stock1[1] #Buying dollars_per_trade$ worth of stock
    units_of_stock2 = 1 / price_stock2[1] #Buying dollars_per_trade$ worth of stock
    
    amount = c(units_of_stock1 * price_stock1[2], units_of_stock2 * price_stock2[2])
    
    sell_what = if (price_stock1[1] / price_stock2[1] > ratio_mean) "A" #stock that is being sold
    else "B"
    profit = if (sell_what == "A") #Calculates profit
            c((1 - amount[1]), (amount[2] - 1), -p * sum(amount))
    else
        c((1 - amount[2]), (amount[1] - 1), -p * sum(amount))
    if (by_stock)
        profit
    else
        sum(profit)
}

model_iterator = 1:floor(nrow(overlap) / 2)
train = overlap[model_iterator, ]
test = overlap[-model_iterator, ]

r.train = train$stock1 / train$stock2 #Setting up the regression model
r.test = test$stock1 / test$stock2 #Setting up the regression model
train.period = seq(min(overlap$Date), by = "5 years", length = 2)#Creates a regression model 

stock1.test = subset(stock1, !(Date >= train.period[1] & Date < train.period[2])) $Adj.Close #Runs regression model
stock2.test = subset(stock2, !(Date >= train.period[1] & Date < train.period[2])) $Adj.Close #Runs regression model
r.test = stock1.test / stock2.test

k.max = max((r.train - mean(r.train)) / sd(r.train))
k.min = min((abs(r.train - mean(r.train)) / sd(r.train)))
ks = seq(k.min, k.max, length = 1000)
m = mean(r.train)

profits = sapply(ks, function(k) { #Creates a list of profits per trade
    pos = get_positions(r.train, k)
    sum(position_profit(pos, train$stock1, train$stock2, mean(r.train)))
})

plot(ks, profits * 100, type = "l", xlab = "k", ylab = "Percent return", main = paste("Profits tested at each std deviation for pairs", name1, "and", name2, sep = " "))
abline(v = k, col = "red")

k.star = mean(ks[profits == max(profits)])
abline(v = k.star, col = 'green')
legend("topleft", legend = c("Chosen std deviation", "Optimal std deviation", "Graph"), fill = c("red", "green", "black"), cex = .60)

print(sprintf("%f$ return in %i trades at a %.2f percent profit", sum(profits), length(profits),sum(profits / 1000) * 100))
```


#Output Chunk
This chunk outputs the return of the top x in this case 10, most and least correlated pairs based on the amount of days, number of stocks, and dollars per trade inputted. This is very similar to the last chunk but instead it's ran and iterated as a function and only gives the outputs. The outputs show that the positively correlated stocks have a much greater return than the negatively correlated ones.
```{r}
trades <- function(days = 1000, num_stocks = 10, dollars_per_trade = 1, index1, index2) {
    files = list.files(path = "./data/", pattern = '\\.csv$', full.names = TRUE)

    for (i in 1: num_stocks) {
        stock1 = read.csv(files[index1[i]]) #Uses the index of the data table instead of the file list
        stock2 = read.csv(files[index2[i]]) #Uses the index of the data table instead of the file list
        
        name1 = substring(files[index1[i]], 8)
        name1 = substr(name1, 1, nchar(name1) - 4) #Stock name without the folder name and file extension
        name2 = substring(files[index2[i]], 8)
        name2 = substr(name2, 1, nchar(name2) - 4) #Stock name without the folder name and file extension

        combine_two_stocks = function(a, b, stockNames = c(deparse(substitute(a)), deparse(substitute(b)))) { #Finds common dates where both stacks were traded 
            rr = range(intersect(a$Date, b$Date))
            a.sub = a[a$Date >= rr[1] & a$Date <= rr[2], ]
            b.sub = b[b$Date >= rr[1] & b$Date <= rr[2], ]
            structure(data.frame(a.sub$Date, a.sub$Adj.Close, b.sub$Adj.Close), names = c("Date", stockNames))
        }
        
        overlap = combine_two_stocks(stock1, stock2)
        r = overlap$stock1 / overlap$stock2
        k = .85 #Standard deviation

        overlap$Date = as.Date(overlap$Date, format = c("%Y-%m-%d")) #Converts overlap$Date into a R readable format

        find_next_position = function(ratio, startDay = 1, k = 1, m = mean(ratio), s = sd(ratio)) { #function that finds which stock to find which stock to buy or sell
            up = m + k * s
            down = m - k * s
            if (startDay > 1)
                ratio = ratio[-(1: (startDay - 1))]
            is_extreme = ratio >= up | ratio <= down
            if (!any(is_extreme))
                return (integer())
            start = which(is_extreme)[1]
            back_to_normal = if (ratio[start] > up)
                    ratio[-(1: start)] <= m
            else
                ratio[-(1: start)] >= m
            end = if (any(back_to_normal))
                    which(back_to_normal)[1] + start
            else
                length(ratio)
            c(start, end) + startDay - 1
        }


        get_positions = function(ratio, k = 1, m = mean(ratio), s = sd(ratio)) { #Finds Where the ratio is optimal to buy or sell
            when = list()
            cur = 1
            while (cur < length(ratio)) {
                temp = find_next_position(ratio, cur, k, m, s)
                if (length(temp) == 0) #quit
                break
                when[[length(when) + 1]] = temp
                if (is.na(temp[2]) || temp[2] == length(ratio))
                    break
                cur = temp[2]
            }
            when
        }
        
        pos = get_positions(r, k) #Finds Where the ratio is optimal to buy or sell

        position_profit = function(pos, stock_price_1, stock_price_2, ratio_mean = mean(stock_price_1 / stock_price_2), p = .001, by_stock = FALSE) { #Finds the profit from the purchases
            if (is.list(pos)) {
                ans = sapply(pos, position_profit,
                    stock_price_1, stock_price_2, ratio_mean, p, by_stock)
                if (by_stock)
                    rownames(ans) = c("A", "B", "commission")
                return (ans)
            }

            price_stock1 = stock_price_1[pos] #Gets prices
            price_stock2 = stock_price_2[pos] #Gets prices

            units_of_stock1 = dollars_per_trade / price_stock1[1] #Buying dollars_per_trade$ worth of stock
            units_of_stock2 = dollars_per_trade / price_stock2[1] #Buying dollars_per_trade$ worth of stock

            amount = c(units_of_stock1 * price_stock1[2], units_of_stock2 * price_stock2[2])
            sell_what = if (price_stock1[1] / price_stock2[1] > ratio_mean) "A" #stock that is being sold
            else "B"
            profit = if (sell_what == "A")
                    c((1 - amount[1]), (amount[2] - 1), -p * sum(amount))
            else
                c((1 - amount[2]), (amount[1] - 1), -p * sum(amount))
            if (by_stock)
                profit
            else
                sum(profit)
        }

        model_iterator = 1: floor(nrow(overlap) / 2)
        train = overlap[model_iterator, ]
        test = overlap[-model_iterator, ]

        r.train = train$stock1 / train$stock2
        r.test = test$stock1 / test$stock2
        train.period = seq(min(overlap$Date), by = "5 years", length = 2) #Creates regression model

        stock1.test = subset(stock1, !(Date >= train.period[1] & Date < train.period[2])) $Adj.Close #Runs regression model
        stock2.test = subset(stock2, !(Date >= train.period[1] & Date < train.period[2])) $Adj.Close #Runs regression model
        r.test = stock1.test / stock2.test

        k.max = max((r.train - mean(r.train)) / sd(r.train))
        k.min = min((abs(r.train - mean(r.train)) / sd(r.train)))

        ks = seq(k.min, k.max, length = days)
        m = mean(r.train)

        profits = sapply(ks, function(k) { #Creates a list of profits per trade
            pos = get_positions(r.train, k)
            sum(position_profit(pos, train$stock1, train$stock2, mean(r.train)))
        })

        cat(sprintf("This returned %.2f dollars in %i days between %s and %s which is an %.2f percent profit using %i dollars per trade\n", sum(profits), length(profits), name1, name2, (sum(profits / days) * 100),dollars_per_trade))
    }
}


print("Positively Correlated")
trades(1000, 10, 1, unique_ordered_correlation$index1, unique_ordered_correlation$index2)
print("Negatively Correlated")
trades(1000, 10, 1, reverse_unique_ordered_correlation$index1, reverse_unique_ordered_correlation$index2)


```